An open reproducible framework for the study of the iterated prisoner's dilemma

### 2016-04-01 tex/paper.tex
"Using jors style file."
This reverts commit 1905b29443dc541fa16dedd80d2c20ec0b09f9b9.

Conflicts:
	tex/paper.pdf
Complete jors styling.

\documentclass{article}

@@

% Lazy page formating (we can adjust when submitting)
\usepackage{fullpage}
\usepackage[parfill]{parskip}
%-----------------------------

\usepackage{amsmath}  % Basic
\usepackage{hyperref}  % Links etc
\usepackage{booktabs}  % Better tables
\usepackage{minted}  % Code
\usepackage{graphicx}  % Images
\usepackage{subcaption}  % Subfigures
\usepackage{authblk}  % Authors

\usepackage{multicol}

\title{An open reproducible framework for the study of the iterated prisoner's
dilemma}

% Core devs (alphabetic order of last name)
\author[1]{Owen Campbell}
\author[1]{Mark Harper}
\author[2]{Vincent Knight}
\author[1]{Karol Langner}

% All devs (alphabetic order of last name)
\author[2]{James Campbell}
\author[5]{Thomas Campbell}
\author[2]{Alex Carney}
\author[2]{Martin Chorley}
\author[1]{Cameron Davidson-Pilon}
\author[1]{Kristian Glass}
\author[1]{Tom{\'a}{\v s} Ehrlich}
\author[1]{Martin Jones}
\author[4]{Georgios Koutsovoulos}
\author[1]{Holly Marissa}
\author[1]{Jochen M{\"u}ller}
\author[2]{Geraint Palmer}
\author[3]{Paul Slavin}
\author[2]{Timothy Standen}
\author[1]{Luis Visintini}
\author[1]{Karl Molden}

% Affiliations
\affil[1]{Not affiliated}
\affil[2]{Cardiff University}
\affil[3]{The University of Manchester}
\affil[4]{The University of Edinburgh}
\affil[5]{St. Nicholas Catholic High School, Hartford}

% Corresponding author

\date{\today}

\begin{document}

\maketitle
\begin{center}
    \thanks{Corresponding author: Vincent Knight
        \href{mailto:knightva@cf.ac.uk}{knightva@cf.ac.uk}}
\end{center}

\begin{abstract}
\noindent
    The Axelrod library is an open source Python package that allows for
    reproducible game theoretic research into the Iterated Prisoner's Dilemma.
    This area of research began in the 1980s but suffers from a lack of
    documentation and test code. The goal of the library is to provide such
    a resource, with facilities for the design of new strategies and
    interactions between them, as well as conducting tournaments and ecological
    simulations for populations of strategies.

\noindent
    With a growing collection of 120 strategies, the library is a also a
    platform for an original tournament that, in itself, is of interest to the
    game theoretic community.

\noindent
    This paper describes the Iterated Prisoner's Dilemma, the Axelrod library
    and its development, and insights gained from some novel research.
\end{abstract}

\section{Overview}\label{sec:sec:overview}

@@

\begin{itemize}
    \item To enable the reproduction of Iterated Prisoner's Dilemma
    research as easily as possible
    \item To produce the de-facto tool for any future Iterated Prisoner's
    Dilemma research
    \item To provide as simple a means as possible for anyone to define and
    contribute new and original Iterated Prisoner's Dilemma strategies
\end{itemize}

@@

\begin{itemize}
    \item Open: all code is released under an MIT license;
    \item Reproducible and well-tested: at the time of writing there is an excellent level of
        integrated tests with 99.37\% coverage
    \item Well-documented: all features of the library are documented for ease of
        use and modification
    \item Extensive: 123 strategies are included, with infinitely-many
        strategies available in the case of parametrised strategies
    \item Extensible: easy to modify to include new strategies and to run new tournaments
\end{itemize}

Before describing the package in more detail in
Section~\ref{sec:description-of-axelrod-python}, an overview of some previous
Iterated Prisoner's Dilemma research will be given in
Section~\ref{sec:review}. Section~\ref{sec:new-strategies-and-implications} will
describe some early research and a conclusion is offered in
Section~\ref{sec:conclusion}.

\subsection{Review of the literature}\label{sec:review}

@@

The tournament is based on an iterated game (see~\cite{Maschler2013} or similar
for details) where two players repeatedly play the normal form game of
(\ref{equ:one-shot}) in full knowledge of each other's playing history to date.
An excellent description of the \textit{one shot} game is given
in~\cite{Gotts2003} which is paraphrased below:

@@

\begin{itemize}
    \item If both choose \(C\), they receive a payoff of \(R\)
        (\textbf{R}eward);
    \item If both choose \(D\), they receive a payoff of \(P\)
        (\textbf{P}unishment);
    \item If one chooses \(C\) and the other \(D\), the defector receives a
        payoff of \(T\) (\textbf{T}emptation) and the cooperator a payoff of
        \(S\) (\textbf{S}ucker).
\end{itemize}

@@

The game of (\ref{equ:one-shot}) is called the Prisoner's Dilemma. Specific
numerical values of \((R,S,T,P)=(3,0,5,1)\) are often used in the literature,
although any satisfying the conditions in \ref{equ:one-shot} will yield similar
results.  Axelrod's tournaments (and further implementations of these) are
sometimes referred to as Iterated Prisoner's Dilemma (IPD) tournaments. An
incomplete overview of published tournaments is given in
Table~\ref{tab:tournaments}.

@@

\section{Implementation and architecture}

\subsection{Description of the Axelrod Python package}\label{sec:description-of-axelrod-python}

@@

\begin{itemize}
    \item Machine learning \cite{scikit-learn} (\url{http://scikit-learn.org/});
    \item Visualisation \cite{Hunter:2007} (\url{http://matplotlib.org/});
    \item Mathematics \cite{sage} (\url{http://www.sagemath.org/});
    \item Astrophysics \cite{astropy} (\url{http://www.astropy.org/});
    \item Data manipulation \cite{mckinney-proc-scipy-2010} (\url{http://pandas.pydata.org/});
    \item Algorithmic Game Theory \cite{Mckelvey06gambit:software} (\url{http://gambit.sourceforge.net/})).
\end{itemize}

@@

As stated in Section~\ref{sec:sec:overview}, one of the main goals of the library
is to allow for the easy contribution of strategies. Doing this requires the
writing of a simple Python class (which can inherit from other predefined
classes). Full contribution guidelines can be found in the documentation.
As an example, Figures~\ref{fig:grudger} and~\ref{fig:grudger_test} show the source code for
the Grudger strategy as well as its corresponding test.

@@

Section~\ref{sec:new-strategies-and-implications} will describe in more detail
some of research capabilities of the library.

\subsection{New strategies, tournaments and implications}\label{sec:new-strategies-and-implications}

@@

\begin{multicols}{3}
    \begin{enumerate}
        \item Cooperator
        \item Defector
        \item ZD-Extort-2
        \item Joss: 0.9
        \item Hard Tit For Tat
        \item Hard Tit For 2 Tats
        \item Tit For Tat
        \item Grudger
        \item Tit For 2 Tats
        \item Win-Stay Lose-Shift
        \item Random: 0.5
        \item ZD-GTFT-2
        \item GTFT: 0.33
        \item Hard Prober
        \item Prober
        \item Prober 2
        \item Prober 3
        \item Calculator
        \item Hard Go By Majority
    \end{enumerate}
\end{multicols}

@@

\begin{itemize}
    \item A closer look at zero determinant strategies, showing that
        extortionate strategies obtain a large number of wins \textit{but} do
        not perform particularly well. This is relevant given the findings of
        \cite{Stewart2012} in which zero determinant strategies are shown to be
        able to perform better than any other strategy. This finding extends
        to noisy tournaments (which are also implemented in the library).
    \item This negative relationship between wins and performance does not
        generalise. There are some strategies that perform well, both in terms
        of matches won and overall performance: Back stabber, Double crosser,
        Looker Up, and Fool Me Once. These strategies continue to perform well in noisy
        tournaments, however some of these have knowledge of the length of the
        game (Back stabber and Double crosser). This is not necessary to rank
        well in both wins and score as demonstrated by Looker Up and Fool Me
        Once.
    \item Strategies like Looker Up and Meta Hunter seem to be generally
        cooperative yet still exploit naive strategies. The MetaHunter strategy
        is a particular type of Meta strategy which uses a variety of other
        strategy behaviours to choose a best action. These strategies perform
        very well in general and continue to do so in noisy tournaments.
\end{itemize}

@@

\section{Availability}\label{sec:availability}

\subsection{Operating system}

@@

\subsection{Programming language}

@@

\subsection{Additional system requirements}

@@

\subsection{Support}

@@

\begin{itemize}
    \item An online chat channel:
        \url{https://gitter.im/Axelrod-Python/Axelrod}.
    \item An  email group:
        \url{https://groups.google.com/forum/#!topic/axelrod-python}.
\end{itemize}

\subsection{Dependencies}

@@

\begin{multicols}{3}
    \begin{itemize}
        \item Numpy 1.9.2:
        \item Matplotlib 1.4.2:
        \item Cloudpickle  0.1.1:
        \item Hypothesis  3.0:
        \item Testfixtures  4.9.1:
    \end{itemize}
\end{multicols}

\subsection{List of contributors}

@@

\begin{multicols}{3}
    \begin{itemize}
        \item {Owen Campbell}
        \item {Mark Harper}
        \item {Vincent Knight}
        \item {Karol Langner}
        \item {James Campbell}
        \item {Thomas Campbell}
        \item {Alex Carney}
        \item {Martin Chorley}
        \item {Cameron Davidson-Pilon}
        \item {Kristian Glass}
        \item {Tom{\'a}{\v s} Ehrlich}
        \item {Martin Jones}
        \item {Georgios Koutsovoulos}
        \item {Holly Marissa}
        \item {Jochen M{\"u}ller}
        \item {Geraint Palmer}
        \item {Paul Slavin}
        \item {Timothy Standen}
        \item {Luis Visintini}
        \item {Karl Molden}
        \item {Jason Young}
        \item {Andy Boot}
        \item {Anna Barriscale}
    \end{itemize}
\end{multicols}

\subsection{Archive}

\begin{multicols}{2}
    \begin{itemize}
        \item Name: Zenodo
        \item Persistent identifier:
\href{https://zenodo.org/record/45187}{10.5281/zenodo.45187}
        \item Licence: MIT
        \item Publisher:  Vincent Knight
        \item Version published: Axelrod: Alternator Release. v0.0.28
        \item Date published: 2016-01-26
    \end{itemize}
\end{multicols}

\subsection{Code repository}

\begin{multicols}{2}
    \begin{itemize}
        \item Name: Github
        \item Identifier: \url{https://github.com/Axelrod-Python/Axelrod}
        \item Licence: MIT
        \item Date published: 2015-02-16
    \end{itemize}
\end{multicols}

\section{Reuse potential}\label{sec:reuse}

@@

This has been evidenced by the work described in
Section~\ref{sec:new-strategies-and-implications} the library has already shown
it's potential for reuse.

\section{Conclusion}\label{sec:conclusion}

@@

\begin{itemize}
    \item Noisy tournaments;
    \item Tournaments with probabilistic ending of interactions;
    \item Ecological analysis of tournaments;
    \item Morality metrics based on \cite{Singer-Clark2014};
    \item Transformation of strategies (in effect giving an infinite number of
        strategies);
    \item Classification of strategies according to multiple dimensions.
    \item Gathering of full interaction history for all interactions.
\end{itemize}

>>>>

\documentclass{jors}

%% Set the header information

% Commenting out as this overrides the section titles.
%\pagestyle{fancy}
%\definecolor{mygray}{gray}{0.6}
%\renewcommand\headrule{}
%\rhead{\footnotesize 3}
%\rhead{\textcolor{gray}{UP JORS software Latex paper template version 0.1}}

@@

\usepackage{amsmath}  % Basic
\usepackage{booktabs}  % Better tables
\usepackage{graphicx}  % Images
\usepackage{hyperref}  % Links etc
\usepackage{minted}  % Code
\usepackage{multicol}  % Multi column list
\usepackage{subcaption}  % Subfigures

\begin{document}
%% Set the header information
\pagestyle{fancy}
\definecolor{mygray}{gray}{0.6}
\renewcommand\headrule{}
\rhead{\footnotesize 3}
\rhead{\textcolor{gray}{UP JORS software Latex paper template version 0.1}}

\rule{\textwidth}{1pt}

\section*{(1) Overview}\label{sec:sec:overview}

\vspace{0.5cm}

\section*{Title}

An open reproducible framework for the study of the iterated prisoner's
dilemma.

\section*{Authors}

\begin{multicols}{2}
    \begin{enumerate}[noitemsep,topsep=0pt]
        % Core devs (alphabetic order of last name)
        \item Campbell, Owen
        \item Harper, Marc
        \item Knight, Vincent
        \item Langner, Karol
        % All devs (alphabetic order of last name)
        \item Campbell, James
        \item Campbell, Thomas
        \item Carney, Alex
        \item Chorley, Martin
        \item Davidson-Pilon, Cameron
        \item Glass, Kristian
        \item Ehrlich Tom{\'a}{\v s}
        \item Jones, Martin
        \item Koutsovoulos, Georgios
        \item Marissa, Holly
        \item M{\"u}ller, Jochen
        \item Palmer, Geraint
        \item Slavin, Paul
        \item Standen, Timothy
        \item Visintini, Luis
        \item Molden, Karl
    \end{enumerate}
\end{multicols}

\section*{Paper Author Roles and Affiliations}

\begin{multicols}{2}
    \begin{enumerate}[noitemsep,topsep=0pt]
\item Development; Not affiliated
\item Development; Not affiliated
\item Development; Knight, Vincent
\item Development; Cardiff University
\item Development; Cardiff University
\item Development; St. Nicholas Catholic High School, Hartford
\item Development; Cardiff University
\item Development; Cardiff University
\item Development; Not affiliated
\item Development; Not affiliated
\item Development; Not affiliated
\item Development; Not affiliated
\item Development; The University of Edinburgh
\item Development; Not affiliated
\item Development; Not affiliated
\item Development; Not affiliated
\item Development; The University of Manchester
\item Development; Cardiff University
\item Development; Not affiliated
\item Development; Not affiliated
    \end{enumerate}
\end{multicols}

\section*{Abstract}

The Axelrod library is an open source Python package that allows for
reproducible game theoretic research into the Iterated Prisoner's Dilemma.
This area of research began in the 1980s but suffers from a lack of
documentation and test code. The goal of the library is to provide such
a resource, with facilities for the design of new strategies and
interactions between them, as well as conducting tournaments and ecological
simulations for populations of strategies.

With a growing collection of 120 strategies, the library is a also a
platform for an original tournament that, in itself, is of interest to the
game theoretic community.

This paper describes the Iterated Prisoner's Dilemma, the Axelrod library
and its development, and insights gained from some novel research.

\section*{Keywords}

Game Theory; Prisoners Dilemma; Python

\section*{Introduction}

@@

\begin{itemize}[noitemsep,topsep=0pt]
    \item To enable the reproduction of Iterated Prisoner's Dilemma
    research as easily as possible
    \item To produce the de-facto tool for any future Iterated Prisoner's
    Dilemma research
    \item To provide as simple a means as possible for anyone to define and
    contribute new and original Iterated Prisoner's Dilemma strategies
\end{itemize}

@@

\begin{itemize}[noitemsep,topsep=0pt]
    \item Open: all code is released under an MIT license;
    \item Reproducible and well-tested: at the time of writing there is an excellent level of
        integrated tests with 99.37\% coverage
    \item Well-documented: all features of the library are documented for ease of
        use and modification
    \item Extensive: 123 strategies are included, with infinitely-many
        strategies available in the case of parametrised strategies
    \item Extensible: easy to modify to include new strategies and to run new tournaments
\end{itemize}

\section*{Review of the literature}\label{sec:review}

@@

The tournament is based on an iterated game (see~\cite{Maschler2013} or similar
for details) where two players repeatedly play the normal form game
of~(\ref{equ:one-shot}) in full knowledge of each other's playing history to
date.  An excellent description of the \textit{one shot} game is given
in~\cite{Gotts2003} which is paraphrased below:

@@

\begin{itemize}[noitemsep,topsep=0pt]
    \item If both choose \(C\), they receive a payoff of \(R\)
        (\textbf{R}eward);
    \item If both choose \(D\), they receive a payoff of \(P\)
        (\textbf{P}unishment);
    \item If one chooses \(C\) and the other \(D\), the defector receives a
        payoff of \(T\) (\textbf{T}emptation) and the cooperator a payoff of
        \(S\) (\textbf{S}ucker).
\end{itemize}

@@

The game of~(\ref{equ:one-shot}) is called the Prisoner's Dilemma. Specific
numerical values of \((R,S,T,P)=(3,0,5,1)\) are often used in the literature,
although any satisfying the conditions in~\ref{equ:one-shot} will yield similar
results.  Axelrod's tournaments (and further implementations of these) are
sometimes referred to as Iterated Prisoner's Dilemma (IPD) tournaments. An
incomplete overview of published tournaments is given in
Table~\ref{tab:tournaments}.

@@

\section*{Implementation and architecture}

\section*{Description of the Axelrod Python package}\label{sec:description-of-axelrod-python}

@@

\begin{itemize}[noitemsep,topsep=0pt]
    \item Algorithmic Game Theory \cite{Mckelvey06gambit:software}
          (\url{http://gambit.sourceforge.net/})).
    \item Astrophysics \cite{astropy} (\url{http://www.astropy.org/});
    \item Data manipulation \cite{mckinney-proc-scipy-2010}
          (\url{http://pandas.pydata.org/});
    \item Machine learning \cite{scikit-learn} (\url{http://scikit-learn.org/});
    \item Mathematics \cite{sage} (\url{http://www.sagemath.org/});
    \item Visualisation \cite{Hunter:2007} (\url{http://matplotlib.org/});
\end{itemize}

@@

As stated in the Overview, one of the main goals of the library
is to allow for the easy contribution of strategies. Doing this requires the
writing of a simple Python class (which can inherit from other predefined
classes). Full contribution guidelines can be found in the documentation.
As an example, Figures~\ref{fig:grudger} and~\ref{fig:grudger_test} show the source code for
the Grudger strategy as well as its corresponding test.

@@

\section*{New strategies, tournaments and implications}\label{sec:new-strategies-and-implications}

@@

\begin{multicols}{2}
    \begin{enumerate}[noitemsep,topsep=0pt]
        \item Cooperator
        \item Defector
        \item ZD-Extort-2
        \item Joss: 0.9
        \item Hard Tit For Tat
        \item Hard Tit For 2 Tats
        \item Tit For Tat
        \item Grudger
        \item Tit For 2 Tats
        \item Win-Stay Lose-Shift
        \item Random: 0.5
        \item ZD-GTFT-2
        \item GTFT: 0.33
        \item Hard Prober
        \item Prober
        \item Prober 2
        \item Prober 3
        \item Calculator
        \item Hard Go By Majority
    \end{enumerate}
\end{multicols}

@@

\begin{itemize}[noitemsep,topsep=0pt]
    \item A closer look at zero determinant strategies, showing that
        extortionate strategies obtain a large number of wins \textit{but} do
        not perform particularly well. This is relevant given the findings of
        \cite{Stewart2012} in which zero determinant strategies are shown to be
        able to perform better than any other strategy. This finding extends
        to noisy tournaments (which are also implemented in the library).
    \item This negative relationship between wins and performance does not
        generalise. There are some strategies that perform well, both in terms
        of matches won and overall performance: Back stabber, Double crosser,
        Looker Up, and Fool Me Once. These strategies continue to perform well in noisy
        tournaments, however some of these have knowledge of the length of the
        game (Back stabber and Double crosser). This is not necessary to rank
        well in both wins and score as demonstrated by Looker Up and Fool Me
        Once.
    \item Strategies like Looker Up and Meta Hunter seem to be generally
        cooperative yet still exploit naive strategies. The MetaHunter strategy
        is a particular type of Meta strategy which uses a variety of other
        strategy behaviours to choose a best action. These strategies perform
        very well in general and continue to do so in noisy tournaments.
\end{itemize}

@@

\section*{(2) Availability}\label{sec:availability}

\section*{Operating system}

@@

\section*{Programming language}

@@

\section*{Additional system requirements}

@@

\section*{Support}

@@

\begin{itemize}[noitemsep,topsep=0pt]
    \item An online chat channel:
        \url{https://gitter.im/Axelrod-Python/Axelrod}.
    \item An  email group:
        \url{https://groups.google.com/forum/#!topic/axelrod-python}.
\end{itemize}

\section*{Dependencies}

@@

\begin{multicols}{3}
    \begin{itemize}[noitemsep,topsep=0pt]
        \item Numpy 1.9.2:
        \item Matplotlib 1.4.2:
        \item Cloudpickle  0.1.1:
        \item Hypothesis  3.0:
        \item Testfixtures  4.9.1:
    \end{itemize}
\end{multicols}

\section*{List of contributors}

@@

\begin{multicols}{2}
    \begin{itemize}[noitemsep,topsep=0pt]
        \item {Owen Campbell}
        \item {Mark Harper}
        \item {Vincent Knight}
        \item {Karol Langner}
        \item {James Campbell}
        \item {Thomas Campbell}
        \item {Alex Carney}
        \item {Martin Chorley}
        \item {Cameron Davidson-Pilon}
        \item {Kristian Glass}
        \item {Tom{\'a}{\v s} Ehrlich}
        \item {Martin Jones}
        \item {Georgios Koutsovoulos}
        \item {Holly Marissa}
        \item {Jochen M{\"u}ller}
        \item {Geraint Palmer}
        \item {Paul Slavin}
        \item {Timothy Standen}
        \item {Luis Visintini}
        \item {Karl Molden}
        \item {Jason Young}
        \item {Andy Boot}
        \item {Anna Barriscale}
    \end{itemize}
\end{multicols}

\section*{Software location:}

{\bf Archive}

\begin{description}[noitemsep,topsep=0pt]
    \item[Name:] Zenodo
    \item[Persistent identifier:]
\href{https://zenodo.org/record/45187}{10.5281/zenodo.45187}
    \item[Licence:] MIT
    \item[Publisher:]  Vincent Knight
    \item[Version published:] Axelrod: Alternator Release. v0.0.28
    \item[Date published:] 2016-01-26
\end{description}

{\bf Code repository}

\begin{description}[noitemsep,topsep=0pt]
    \item[Name:] Github
    \item[Identifier:] \url{https://github.com/Axelrod-Python/Axelrod}
    \item[Licence:] MIT
    \item[Date published:] 2015-02-16
\end{description}

\section*{Reuse potential}\label{sec:reuse}

@@

This has been evidenced by the work described: the library has already shown
it's potential for reuse.

\section*{Conclusion}\label{sec:conclusion}

@@

\begin{itemize}[noitemsep,topsep=0pt]
    \item Noisy tournaments;
    \item Tournaments with probabilistic ending of interactions;
    \item Ecological analysis of tournaments;
    \item Morality metrics based on \cite{Singer-Clark2014};
    \item Transformation of strategies (in effect giving an infinite number of
        strategies);
    \item Classification of strategies according to multiple dimensions.
    \item Gathering of full interaction history for all interactions.
\end{itemize}
